Trajectories

### 2021-09-20 Rmd/report.Rmd
Add LOO

# import data
tar_load(items)
tar_load(participants)
tar_load(vocabulary)
tar_load(responses)
tar_load(fits_comp)
tar_load(fits_prod)

@@

```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
a <- participants %>% 
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  count(age, doe_2) %>% 
  arrange(doe_2) %>% 
  mutate(doe_2 = fct_inorder(doe_2)) %>% 
  pivot_wider(names_from = doe_2, values_from = n, values_fill = 0) %>% 
  gt() %>% 
  tab_spanner(label = "Number of participants by DOE", columns = 2:6) %>% 
  data_color(
    columns = 2:6,
    colors = scales::col_numeric(
      palette = c("white", "orange"),
      domain = c(0, 11)
    )
  ) %>% 
  cols_label(
    age = md("**Age<br>(months)**")
  ) %>% 
  as_raw_html()

@@

@@

### Short vs. Long {.tabset .tab-pills}

```{r vocabulary_long_comprehension, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary$long %>% 
  filter(type=="understands") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point() +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

```{r vocabulary_long_production, message=FALSE, warning=FALSE, paged.print=FALSE}
vocabulary$long %>% 
  filter(type=="understands") %>% 
  ggplot(aes(value, reorder(id, age), colour = version, fill = version, shape = version)) +
  facet_grid(~scale) +
  geom_point(size = 3) +
  labs(x= "Vocabulary size", y = "Participant (ordered by age, min-to-max)",
       colour = "Version", fill = "Version", shape = "Version") +
  scale_x_continuous(labels = scales::percent) +
  theme_github() +
  theme(
    panel.grid.major.y = element_blank(),
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank()
  )
```

@@

```{r raw, message=FALSE, warning=FALSE, paged.print=FALSE}
responses %>%
  group_by(te, cognate) %>%
  summarise(
    understands = mean(understands, na.rm = TRUE),
    produces = mean(produces, na.rm = TRUE)
  ) %>%
  pivot_longer(
    c(understands, produces),
    names_to = "type",
    values_to = "prop"
  ) %>%
  mutate(type = str_to_sentence(type)) %>% 
  ggplot(aes(prop, reorder(te, -prop), colour = cognate)) +
  facet_wrap(vars(type)) +
  geom_vline(xintercept = 0.5, linetype = "dotted", size = 1) +
  geom_point(alpha = 0.5, size = 2) +
  labs(x = "Mean % of acquisition", y = "Translation equivalent ID", colour = "Type") +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_github() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
```

```{r raw_age, message=FALSE, warning=FALSE, paged.print=FALSE}
responses %>%
  mutate(age = as.factor(round(age))) %>% 
  group_by(te, age, cognate) %>%
  summarise(
    understands = mean(understands, na.rm = TRUE),
    produces = mean(produces, na.rm = TRUE)
  ) %>%
  pivot_longer(
    c(understands, produces),
    names_to = "type",
    values_to = "prop"
  ) %>%
  mutate(type = str_to_sentence(type)) %>% 
  ggplot(aes(age, prop, colour = cognate, fill = cognate, shape = cognate)) +
  facet_wrap(vars(type)) +
  geom_hline(yintercept = 0.5, linetype = "dotted", size = 1) +
  geom_point(aes(group = interaction(age, cognate)), alpha = 0.5, size = 0.5) +
  stat_summary(fun.data = mean_se, geom = "pointrange") +
  labs(x = "Mean % of acquisition", y = "Translation equivalent ID",
       colour = "Cognate", shape = "Cognate", fill = "Cognate") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_github() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    legend.title = element_blank()
  )
```

@@

# emmeans
nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = unique(participants$lp),
  cognate = unique(items$cognate),
  frequency_center = c(-1, 0, 1)
)

m <- add_fitted_draws(nd, fits_comp$fit_2, n = 20, re_formula = NA) %>% 
  mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))

@@

# emmeans
nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = unique(participants$lp),
  cognate = unique(items$cognate),
  frequency_center = c(-1, 0, 1)
)

@@

@@

ggplot(re_id, aes(Estimate, reorder(id, Estimate))) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
                 colour = wes_palettes$Darjeeling1[1]) +
  geom_point(size = 1, shape = "|") +
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  labs(x = "Participant average probability of comprehension",
       y = "Participant") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_github() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )

@@

ggplot(re_te, aes(Estimate, reorder(te, Estimate))) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
                 colour = wes_palettes$Darjeeling1[2]) +
  geom_point(size = 1, shape = "|") +
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  labs(x = "Item average probability of comprehension",
       y = "Item") +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_github() +
  theme(
    axis.text.y = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank()
  )
```

@@

ggplot(re_id, aes(Estimate, reorder(id, Estimate))) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
                 colour = wes_palettes$Darjeeling1[1]) +
  geom_point(size = 1, shape = "|") +
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  labs(x = "Participant average probability of comprehension",
       y = "Participant") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_github() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_blank()
  )

@@

ggplot(re_te, aes(Estimate, reorder(te, Estimate))) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25,
                 colour = wes_palettes$Darjeeling1[2]) +
  geom_point(size = 1, shape = "|") +
  geom_vline(xintercept = 0.5, linetype = "dashed") +
  labs(x = "Item average probability of comprehension",
       y = "Item") +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_github() +
  theme(
    axis.text.y = element_blank(),
    panel.grid = element_blank(),
    axis.ticks = element_blank()
  )
```

@@

nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = unique(participants$lp),
  cognate =unique(items$cognate),
  frequency_center = c(-1,  0, 1)
)

>>>>

# import data
tar_load(items)
tar_load(participants)
tar_load(vocabulary)
tar_load(responses)
tar_load(fits_comp)
tar_load(fits_prod)
tar_load(fits_comp_prior)
tar_load(fits_prod_prior)
tar_load(loo_comp)
tar_load(loo_prod)

@@

```{r participants_lp, message=FALSE, warning=FALSE, paged.print=FALSE}
labs <- c("0-10%", "10-20%", "20-30%", "30-40%", "40-50%")
participants %>% 
  mutate(
    age = round(age),
    doe_2 = ifelse(dominant_language=="Catalan", doe_spanish, doe_catalan),
    doe_2 = cut(doe_2, breaks = seq(0, 0.5, 0.1), include.lowest = TRUE, labels = labs)
  ) %>% 
  count(age, doe_2) %>% 
  arrange(doe_2) %>% 
  mutate(doe_2 = fct_inorder(doe_2)) %>% 
  pivot_wider(names_from = doe_2, values_from = n, values_fill = 0) %>% 
  gt() %>% 
  tab_spanner(label = "Number of participants by DOE", columns = 2:6) %>% 
  data_color(
    columns = 2:6,
    colors = scales::col_numeric(
      palette = c("white", "orange"),
      domain = c(0, 30)
    )
  ) %>% 
  cols_label(
    age = md("**Age<br>(months)**")
  ) %>% 
  as_raw_html()

@@

## SES/parental education

```{r participants_edu, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
participants %>%
  mutate(edu_parent = fct_explicit_na(as.factor(edu_parent))) %>% 
  count(lp, dominant_language, edu_parent) %>% 
  right_join(
    expand_grid(
      lp = c("Monolingual", "Bilingual"),
      dominant_language = c("Catalan", "Spanish"),
      edu_parent = unique(participants$edu_parent)
    )
  ) %>% 
  replace_na(list(n = 0)) %>% 
  pivot_wider(names_from = edu_parent, values_from = n, values_fill = 0) %>%
  mutate(n = rowSums(cbind(.[,3:8]))) %>% 
  select(-"NA") %>% 
  gt() %>% 
  cols_label(
    lp = "Lang. Profile",
    dominant_language = "L1"
  ) %>% 
  summary_rows(
    columns = 3:8,
    fns = list(n = ~sum(.)),
    formatter = fmt_number,
    decimals = 0
  ) %>% 
  tab_style(
    style = list(cell_borders(weight = px(2), sides = "left")),
    locations = cells_body(columns = 10, rows = 1:4)
  )
```

@@

@@

```{r raw, message=FALSE, warning=FALSE, paged.print=FALSE}
responses %>%
  group_by(te, cognate) %>%
  summarise(
    understands = mean(understands, na.rm = TRUE),
    produces = mean(produces, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    c(understands, produces),
    names_to = "type",
    values_to = "prop"
  ) %>%
  mutate(type = str_to_sentence(type)) %>% 
  ggplot(aes(prop, reorder(te, -prop), colour = cognate)) +
  facet_wrap(vars(type)) +
  geom_vline(xintercept = 0.5, linetype = "dotted", size = 1) +
  geom_point(alpha = 0.5, size = 2) +
  labs(x = "Mean % of acquisition", y = "Translation equivalent ID", colour = "Type") +
  scale_x_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_github() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
  )
```

```{r raw_age, message=FALSE, warning=FALSE, paged.print=FALSE}
responses %>%
  mutate(age = as.factor(round(age))) %>% 
  group_by(te, age, cognate) %>%
  summarise(
    understands = mean(understands, na.rm = TRUE),
    produces = mean(produces, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    c(understands, produces),
    names_to = "type",
    values_to = "prop"
  ) %>%
  mutate(type = str_to_sentence(type)) %>% 
  ggplot(aes(age, prop, colour = cognate, fill = cognate, shape = cognate)) +
  facet_wrap(vars(type)) +
  geom_hline(yintercept = 0.5, linetype = "dotted", size = 1) +
  geom_point(aes(group = interaction(age, cognate)), alpha = 0.5, size = 0.5) +
  stat_summary(fun.data = mean_se, geom = "pointrange") +
  labs(x = "Mean % of acquisition", y = "Translation equivalent ID",
       colour = "Cognate", shape = "Cognate", fill = "Cognate") +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  theme_github() +
  theme(
    legend.position = "top",
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank(),
    legend.title = element_blank()
  )
```

@@

# emmeans
nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = c("Monolingual", "Bilingual"),
  cognate = unique(items$cognate),
  frequency_center = c(-1, 0, 1)
)

m <- add_fitted_draws(nd, fits_comp_prior$fit_2, n = 20, re_formula = NA) %>% 
  mutate(frequency_center = as.factor(paste0("Frequency = ", frequency_center, " SD")))

@@

# emmeans
nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = c("Monolingual", "Bilingual"),
  cognate = unique(items$cognate),
  frequency_center = c(-1, 0, 1)
)

@@

### Comprehension

```{r loo_comp, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
loo_comp %>% 
  as_tibble() %>% 
  rownames_to_column("model") %>% 
  relocate(model, matches("elpd_loo"), matches("p_loo"), matches("looic"), matches("diff")) %>%
  mutate_all(as.numeric) %>% 
  gt() %>% 
  tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
  tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
  tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
  tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
  fmt_number(3:13) %>% 
  cols_label(
    model = "Model",
    # elpd
    elpd_loo = md("*ELPD*"),
    se_elpd_loo = md("*SE*"),
    subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
    # p
    p_loo = md("*p*"),
    se_p_loo = md("SE"),
    subsampling_se_p_loo = md("SE<sub>sub</sub>"),
    # looic
    looic = md("*LOO-IC*"),
    se_looic = md("*SE*"),
    subsampling_se_looic = md("SE<sub>sub</sub>"),
    # diff
    elpd_diff = md("*diff*"),
    se_diff = md("*SE*"),
    subsampling_se_diff = md("*SE<sub>sub</sub>*"),
  ) %>% 
  as_raw_html()
```

### Production

```{r loo_prod, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
loo_comp %>% 
  as_tibble() %>% 
  rownames_to_column("model") %>% 
  relocate(model, matches("elpd_loo"), matches("p_loo"), matches("looic"), matches("diff")) %>%
  mutate_all(as.numeric) %>% 
  gt() %>% 
  tab_spanner(md("**LOO<sub>ELPD</sub>**"), matches("elpd_loo")) %>% 
  tab_spanner(md("**LOO<sub>p</sub>**"), matches("p_loo")) %>% 
  tab_spanner(md("**LOO<sub>IC</sub>**"), matches("looic")) %>% 
  tab_spanner(md("**LOO<sub>diff</sub>**"), matches("diff")) %>% 
  fmt_number(3:13) %>% 
  cols_label(
    model = "Model",
    # elpd
    elpd_loo = md("*ELPD*"),
    se_elpd_loo = md("*SE*"),
    subsampling_se_elpd_loo = md("*SE<sub>sub</sub>*"),
    # p
    p_loo = md("*p*"),
    se_p_loo = md("SE"),
    subsampling_se_p_loo = md("SE<sub>sub</sub>"),
    # looic
    looic = md("*LOO-IC*"),
    se_looic = md("*SE*"),
    subsampling_se_looic = md("SE<sub>sub</sub>"),
    # diff
    elpd_diff = md("*diff*"),
    se_diff = md("*SE*"),
    subsampling_se_diff = md("*SE<sub>sub</sub>*"),
  ) %>% 
  as_raw_html()
```

@@

ggplot(re_id, aes(Estimate, reorder(id, Estimate))) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, colour = "white", height = 0) +
  geom_point(size = 0.5, colour = "white") +
  geom_vline(xintercept = 0.5, colour = "white") +
  labs(x = "Participant average probability of comprehension",
       y = "Participant") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_github() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_line(colour = "white", linetype = "dotted"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.line = element_line(colour = "white")
  )
```

@@

ggplot(re_te, aes(Estimate, reorder(te, Estimate))) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, colour = "white", heigth = 0) +
  geom_point(size = 0.5, colour = "white") +
  geom_vline(xintercept = 0.5, colour = "white") +
  labs(x = "Item average probability of comprehension",
       y = "Item") +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_github() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_line(colour = "white", linetype = "dotted"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.line = element_line(colour = "white")
  )
```

@@

ggplot(re_id, aes(Estimate, reorder(id, Estimate))) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, colour = "white", heigth = 0) +
  geom_point(size = 0.5, colour = "white") +
  geom_vline(xintercept = 0.5, colour = "white") +
  labs(x = "Item average probability of production",
       y = "Item") +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_github() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_line(colour = "white", linetype = "dotted"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.line = element_line(colour = "white")
  )

@@

ggplot(re_te, aes(Estimate, reorder(te, Estimate))) +
  geom_errorbarh(aes(xmin = Q2.5, xmax = Q97.5), size = 0.25, colour = "white", heigth = 0) +
  geom_point(size = 0.5, colour = "white") +
  geom_vline(xintercept = 0.5, colour = "white") +
  labs(x = "Item average probability of production",
       y = "Item") +
  scale_color_brewer(palette = "Set1") +
  scale_x_continuous(limits = c(0, 1)) +
  theme_github() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    panel.grid = element_line(colour = "white", linetype = "dotted"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.line = element_line(colour = "white")
  )
```

@@

nd <- expand.grid(
  age = seq(min(responses$age), max(responses$age), length.out = 50),
  lp = c("Monolingual", "Bilingual"),
  cognate = unique(items$cognate),
  frequency_center = c(-1,  0, 1)
)
